On the effects of partying on next-day sports performance

### 2017-06-14 writing/draft.tex
Fixed camel-casing in regression tables.

\begin{center}
\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   logged travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   east-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf. Crit. & 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}
\end{tabular} 
\end{center}

@@

\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of Nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Logged travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf. Crit. & 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}

@@

\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   party placebo & 0.066 \\    & (0.088) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.006 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   number hours rest time & 0.001 \\    & (0.001) \\    & \\   time of game during day & 0.002 \\    & (0.013) \\    & \\   home team effect & $-$0.008 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf. Crit. & 13,631.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}

@@

\paragraph{General}
In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting. 
Future work may consider similar effects in European soccer leagues.

\subsection{Conclusion}
We do find that visiting a city with active nightlife the day before a game does have 
a statistically significant negative impact on a team's likelihood of meeting the spread in NBA. This decreased likelihood to meet the spread may be explained 
by the team's defense letting up more points. In turn, we believe that the drop in
defensive performance may be caused by a drop in physical exertion.
We verify that this phenomena persists across sport, and 
to that end look toward major league baseball money-lines. Using these, we back out gambling
house's estimate for the team's probability of winning the game, and after conditioning on this
we find that visiting a city with active nightlife the day before a game does negatively 
impact the team's likelihood of winning.
With a causal model established, we realize a profitable betting strategy in each year of testing.

>>>>

\begin{center}
\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Log of travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Log of travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf. Crit. & 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}
\end{tabular} 
\end{center}

@@

\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Log of travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf. Crit. & 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}

@@

\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   Party placebo & 0.066 \\    & (0.088) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Log of travel distance & $-$0.006 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   Number hours rest time & 0.001 \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.008 \\    & (0.043) \\    & \\   Log travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf. Crit. & 13,631.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}

@@

\paragraph{General}
In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting; recall that our decision to only look at back-to-back games 
within 24 hours of each other was only motivated by intuition and not guided by data. 
Future work may consider similar effects in European soccer leagues.

\subsection{Conclusions}
We do find that visiting a city with active nightlife the day before a game does have 
a statistically significant negative impact on a team's likelihood of meeting the spread in NBA. This decreased likelihood to meet the spread may be explained 
by the team's defense letting up more points. In turn, we believe that the drop in
defensive performance may be caused by a drop in physical exertion.
We verify that this phenomena persists across sport, and 
to that end look toward major league baseball money-lines. Using these, we back out gambling
house's estimate for the team's probability of winning the game, and after conditioning on this
we find that visiting a city with active nightlife the day before a game does negatively 
impact the team's likelihood of winning.
With a causal model established, we realize a profitable betting strategy in each year of testing.
